---
title: "annpixeltest"
author: "Dominic Cugliari"
date: "2023-06-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath(".."))
```

```{r}
library(tidyverse)
library(readr)
library(neuralnet)
library(Ecdat)
library(keras)
library(Matrix)
library(ggplot2)
library(binhf)
library(OpenImageR)
```



```{r}
ogimg <- read_csv("raw_data/annreference.csv")
ogimg %>% as.matrix()
NX <- 65 # number of horizontal pixels in digital image
NZ <- 65
pixelx = 16 # horizontal pixel position from middle of image 
pixelz = 16 # vertical pixel position from middle of image
Test1_sinogram <- read_csv(
                   'raw_data/annsino.csv')
nangles <- nrow(Test1_sinogram)
ndet <- ncol(Test1_sinogram)
dummy_weight <- tibble(seq(1:ndet) * 0.00e+0)
dummy_weight %>% as.matrix()
dummy_weight[ceiling(ndet/2),] <- 0.25
dummy_weight[ceiling(ndet/2)-1,] <- -.125
dummy_weight[ceiling(ndet/2)+1,] <- -.125
angle_step <- 180/(nangles-1)
dshift <- data.frame( angle = seq(0,nangles-1)*angle_step )
dshift$pixel <- round( pixelx*cos(pi*dshift$angle/180) +                                    pixelz*sin(pi*dshift$angle/180) )
sTest1_sinogram <- Test1_sinogram
```

```{r}
sTest1_sinogram <- sTest1_sinogram %>%
  as.matrix()
sTest1_sinogram <- sTest1_sinogram
sTest1_sinogram
sTest1_sinogram %>% image()

```

sum the columns into a vector

```{r}
for(val in 1:nangles ){
                        sTest1_sinogram[val,] <- shift(sTest1_sinogram[val,],
                                                       dshift$pixel[val], dir = "left")
  }
f_vector <- colSums(sTest1_sinogram)
f_vector
```

```{r}
z_vector <- rev(f_vector)
z_vector
```
Format the z_vector input and output pixel intensity into a data frame and transpose from rows to columns for use in the neuralnet function

```{r}
c_data <- as_tibble(c(z_vector, 1.547052362
))
ANN_data <- t(c_data) # data frame for training ANN in function neuralnet
```

Define names of the columns for the data to used by function neuralnet

```{r}
ANN_names <- c(paste0("z",(as.character(c(seq(1:ndet))))),"o")
colnames(ANN_data) <- ANN_names
```

Define the function f to be used in the neuralnet in terms of the column names of the ANN_data data frame, train the neuralnet and plot the output.

```{r}
f <- as.formula(paste(ANN_names[ndet+1],"~", paste(ANN_names
                  [!ANN_names %in% "o"], collapse = " + ")))

net.FBP <-neuralnet(f, ANN_data, hidden = 2)

plot (net.FBP, rep = "best")

net.FBP

```

```{r}
filt_sinoA <- Test1_sinogram
number_cols <- ncol(Test1_sinogram)

for(val in 1:number_cols){
                    filt_sinoA[,val] <- convolve(Test1_sinogram[,val], net.FBP$startweights[[1]][[1]][2:94,1], type ="circular")
}

filt_sinoA %>% 
  as.matrix()

write_csv(filt_sinoA, "C:/Users/sumloan/Documents/R/anntest/raw_data/filtsino2.csv")

Test1_sinogram %>%
  as.matrix()%>%
  image()
```



EDIT

```{r}
filt_sinoA <- Test1_sinogram
number_cols <- ncol(Test1_sinogram)

for(val in 1:nangles){
                    filt_sinoA[, val] <- convolve(t(Test1_sinogram[val,]), net.FBP$startweights[[1]][[1]][2:94,1], type ="circular")
}

filt_sinoA %>% 
  as.matrix()

write_csv(filt_sinoA, "C:/Users/sumloan/Documents/R/anntest/raw_data/filtsino2.csv")

Test1_sinogram %>%
  as.matrix()%>%
  image()
```















```{r}
annreference <- read_csv("raw_data/annreference.csv")%>%
  as.matrix()
annreference %>% image()


```















































