---
title: "anntest"
author: "Dominic Cugliari"
date: "2023-06-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath(".."))
```

```{r}
library(tidyverse)
library(readr)
library(neuralnet)
library(Ecdat)
library(keras)
```

Load all of the Phantoms

```{r}
img1 <- read.csv("raw_data/Reference1.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
img1
img1 %>% image()

img2 <- read.csv("raw_data/Reference2.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
img2
img2 %>% image()

img3 <- read.csv("raw_data/Reference3.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
img3
img3 %>% image()

img4 <- read.csv("raw_data/Reference4.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
img4
img4 %>% image()

```

Load all of the sinograms

```{r}
sino1 <- read.csv("raw_data/sinogram1.csv")
sino1 <- select(sino1, X..0.707..0.703.:X.0.703.0.707.)
sino1 <- as.matrix(sino)
sino1
image(sino1)


sino2 <- read.csv("raw_data/sinogram2.csv")
sino2 <- select(sino2, X..0.707..0.703.:X.0.703.0.707.)
sino2 <- as.matrix(sino2)
sino2
image(sino2)

sino3 <- read.csv("raw_data/sinogram3.csv")
sino3 <- select(sino3, X..0.707..0.703.:X.0.703.0.707.)
sino3 <- as.matrix(sino3)
sino3
image(sino3)

sino4 <- read.csv("raw_data/sinogram4.csv")
sino4 <- select(sino4, X..0.707..0.703.:X.0.703.0.707.)
sino4 <- as.matrix(sino4)
sino4
image(sino4)
```

Load Filtered Back Projections using Hamming 1

```{r}
FBP1a <- read.csv("raw_data/backProjection1a.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
FBP1a %>% image()

FBP2a <- read.csv("raw_data/backProjection2a.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
FBP2a %>% image()

FBP3a <- read.csv("raw_data/backProjection3a.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
FBP3a %>% image()

FBP4a <- read.csv("raw_data/backProjection4a.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
FBP4a %>% image()
```

Load Filtered Back Projection using Ramp

```{r}
FBP1b <- read.csv("raw_data/backProjection1c.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
FBP1b %>% image()

FBP2b <- read.csv("raw_data/backProjection2c.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
FBP2b %>% image()

FBP3b <- read.csv("raw_data/backProjection3c.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
FBP3b %>% image()

FBP4b <- read.csv("raw_data/backProjection4c.csv")%>%
  select(, V1:V257)%>%
  as.matrix()
FBP4b %>% image()
```
Create the bias matrix

```{r}
bias <- matrix(-1, nrow = 257, ncol = 257)
bias
```


Pick first pixel
```{r}
FBP1av <- FBP1a[1,1]

FBP1bv <- FBP1b[1,1]

img1v <- img1[1,1]

biasv <- bias[1,1]
                                        
```


Create data frame 1
```{r}
test1 <- data.frame(img1v, FBP1av, FBP1bv, biasv)

  

test1
```
Create neural net
```{r}
nn <- neuralnet(formula = img1v ~ FBP1av + FBP1bv + biasv,
          data = test1,
          hidden = 2,
          stepmax = 1e+06,
          )
nn
```
Pull the generalized weights

```{r}
weight1 <- nn$weights
weight1
```

Pick pixel 2
```{r}
FBP1av2 <- FBP1a[2,1]

FBP1bv2 <- FBP1b[2,1]

img1v2 <- img1[2,1]

biasv2 <- bias[2,1]
                                        
```


Create data frame 2
```{r}
test2 <- data.frame(img1v2, FBP1av2, FBP1bv2, biasv2)

  

test2
```
    
Create new neural net
```{r}
nn2 <- neuralnet(formula = img1v2 ~ FBP1av2 + FBP1bv2 + biasv2,
          data = test2,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight1
          )
nn2
```


Pull weight 2

```{r}
weight2 <- nn2$weights
weight2
```

Pick pixel 3
```{r}
FBP1av3 <- FBP1a[3,1]

FBP1bv3 <- FBP1b[3,1]

img1v3 <- img1[3,1]

biasv3 <- bias[3,1]
                                        
```


Create data frame 3
```{r}
test3 <- data.frame(img1v3, FBP1av3, FBP1bv3, biasv3)

  

test3
```
    
Create neural net 3
```{r}
nn3 <- neuralnet(formula = img1v3 ~ FBP1av3 + FBP1bv3 + biasv3,
          data = test3,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight2
          )
nn3
```


Pull weight 3

```{r}
weight3 <- nn3$weights
weight3
```

Pick pixel 4
```{r}
FBP1av4 <- FBP1a[4,1]

FBP1bv4 <- FBP1b[4,1]

img1v4 <- img1[4,1]

biasv4 <- bias[4,1]
                                        
```


Create data frame 4
```{r}
test4 <- data.frame(img1v4, FBP1av4, FBP1bv4, biasv4)

  

test4
```
    
Create new neural net 4
```{r}
nn4 <- neuralnet(formula = img1v4 ~ FBP1av4 + FBP1bv4 + biasv4,
          data = test4,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight3
          )
nn4
```

```{r}
weight1
weight2
weight3
```


Pull weight 4

```{r}
weight4 <- nn4$weights
weight4
```

Pick pixel 5
```{r}
FBP1av5 <- FBP1a[5,1]

FBP1bv5 <- FBP1b[5,1]

img1v5 <- img1[5,1]

biasv5 <- bias[5,1]
                                        
```


Create data frame 5
```{r}
test5 <- data.frame(img1v5, FBP1av5, FBP1bv5, biasv5)

  

test5
```
    
Create new neural net
```{r}
nn5 <- neuralnet(formula = img1v5 ~ FBP1av5 + FBP1bv5 + biasv5,
          data = test5,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight4
          )
nn5
```



Pull weight 5

```{r}
weight5 <- nn5$weights
weight5
```

Pick pixel 6
```{r}
FBP1av6 <- FBP1a[6,1]

FBP1bv6 <- FBP1b[6,1]

img1v6 <- img1[6,1]

biasv6 <- bias[6,1]
                                        
```


Create data frame 6
```{r}
test6 <- data.frame(img1v6, FBP1av6, FBP1bv6, biasv6)

  

test6
```
    
Create new neural net
```{r}
nn6 <- neuralnet(formula = img1v6 ~ FBP1av6 + FBP1bv6 + biasv6,
          data = test6,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight5
          )
nn6
```

Pull weight 6

```{r}
weight6 <- nn6$weights
weight6
```

Pick pixel 7
```{r}
FBP1av7 <- FBP1a[7,1]

FBP1bv7 <- FBP1b[7,1]

img1v7 <- img1[7,1]

biasv7 <- bias[7,1]
                                        
```


Create data frame 7
```{r}
test7 <- data.frame(img1v7, FBP1av7, FBP1bv7, biasv7)

  

test7
```
    
Create new neural net
```{r}
nn7 <- neuralnet(formula = img1v7 ~ FBP1av7 + FBP1bv7 + biasv7,
          data = test7,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight6
          )
nn7
```


Pull weight 7
```{r}
weight7 <- nn7$weights
weight7
```

Pick pixel 8
```{r}
FBP1av8 <- FBP1a[8,1]

FBP1bv8 <- FBP1b[8,1]

img1v8 <- img1[8,1]

biasv8 <- bias[8,1]
                                        
```


Create data frame 8
```{r}
test8 <- data.frame(img1v8, FBP1av8, FBP1bv8, biasv8)

  

test8
```
    
Create new neural net
```{r}
nn8 <- neuralnet(formula = img1v8 ~ FBP1av8 + FBP1bv8 + biasv8,
          data = test8,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight7
          )
nn8
```


Pull weight 8
```{r}
weight8 <- nn8$weights
weight8
```


Pick pixel 9
```{r}
FBP1av9 <- FBP1a[9,1]

FBP1bv9 <- FBP1b[9,1]

img1v9 <- img1[9,1]

biasv9 <- bias[9,1]
                                        
```


Create data frame 9
```{r}
test9 <- data.frame(img1v9, FBP1av9, FBP1bv9, biasv9)

  

test9
```
    
Create new neural net
```{r}
nn9 <- neuralnet(formula = img1v9 ~ FBP1av9 + FBP1bv9 + biasv9,
          data = test9,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight8
          )
nn9
```


Pull weight 9
```{r}
weight9 <- nn9$weights
weight9
```


Pick pixel 10
```{r}
FBP1av10 <- FBP1a[10,1]

FBP1bv10 <- FBP1b[10,1]

img1v10 <- img1[10,1]

biasv10 <- bias[10,1]
                                        
```


Create data frame 10
```{r}
test10 <- data.frame(img1v10, FBP1av10, FBP1bv10, biasv10)

  

test10
```
    
Create new neural net
```{r}
nn10 <- neuralnet(formula = img1v10 ~ FBP1av10 + FBP1bv10 + biasv10,
          data = test10,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight9
          )
nn10
```


Pull weight 10
```{r}
weight10 <- nn10$weights
weight10
```


Pick pixel 11
```{r}
FBP1av11 <- FBP1a[11,1]

FBP1bv11 <- FBP1b[11,1]

img1v11 <- img1[11,1]

biasv11 <- bias[11,1]
                                        
```


Create data frame 10
```{r}
test11 <- data.frame(img1v11, FBP1av11, FBP1bv11, biasv11)

  

test11
```
    
Create new neural net
```{r}
nn11 <- neuralnet(formula = img1v11 ~ FBP1av11 + FBP1bv11 + biasv11,
          data = test11,
          hidden = 2,
          stepmax = 1e+06,
          startweights = weight10
          )
nn11
```


Pull weight 10
```{r}
weight11 <- nn11$weights
weight11
```











