---
title: "CST_Modeling"
author: "Jeff Martin"
date: "2023-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath(".."))
```

Import the tidyverse, Matrix and ggplot2 packages

```{r}
library(tidymodels)
library(Matrix)
library(ggplot2)
library(readr)
```

Define size of image plane

```{r}
NX <- 64 # number of pixels in x direction
NZ <- NX # number of pixels in z direction, square space for now
minx <- 0 # minimum x position in cm
minz <- 0 # minimum y position in cm
maxx <- 12.8  # maximum x position in cm
maxz <- maxx  # maximum z position in cm
pwidth <- maxx/NX # define pixel width in cm
```

Define the width of the slice of the object to be imaged in y direction

```{r}
wd <- 0.254 # width of inspection volume in z direction
wColl <- 0.254 # width of detector collimator slit in z direction
```

Define gamma-ray source characteristics

```{r}
E0 <- 88.04 # Energy of primary gamma-ray of Cd-109 radioisotope in keV
moc2 <- 511 # Rest mass energy of electron in keV from NIST                                                    reference--https://physics.nist.gov/cgi-bin/cuu/Value?mec2mev
S_Act_mCi <- 75 # Input Cd-109 source activity in mCi
mCi_to_Bq <- 37000000 # Conversion factor from mCi to Bq (decays/sec)
S_Act <- mCi_to_Bq*S_Act_mCi # Calculate source activity in Bq
Ints_88k <- 0.0364 # Intensity of 88 keV gamma-rays in # per decay
Em_rate <- Ints_88k*S_Act # Emission rate of 88 keV gamma rays from Cd-109 source
r02 <- 0.0794 #classical electron radius squared in barns (1e-24 cm^2) from NIST                                reference--https://physics.nist.gov/cgi-bin/cuu/Value?re
```

Define gamma-ray source geometry

```{r}
NS <- 8 # number of source (and first detector) locations
XSspace <- 0.5
ZSspace <- 0.1
SourceLoc <- tibble( X = seq( 0, NS-1 ) * XSspace ) #x coordinate of source positions
SourceLoc$Z <- SourceLoc$X * ZSspace/XSspace #z coordinate of source positions

SourceLoc$Theta <- SourceLoc$Z #Initialize vector in SourceLoc for source angles
SourceLoc$Theta[ seq(1, NS) ] <- 29.3 #Angle of source positions in degrees
SourceLoc$Theta[ seq(NS/2+1, NS)] <- 180 - 29.3   #Angle of last 4 source positions in degrees
SourceDX <- 0.508 # source diameter in X dimension in cm
SourceDZ <- 0.1 # source thickness in Z dimension in cm
SourceNX <- 10 # number of source sub-positions in X dimension
SourceNZ<- 2 # number of source sub-positions in Z dimension
```

Define number of detectors and their positions.  For initial calculations space the detectors evenly along the x-axis with the source

```{r}
ndet <- 12 # number of gamma-ray detectors in array
det_dist <- 1.0e+0 # separation in cm of the first detector from the source and between the remaining detectors in the array
det_x <- t(tibble ( seq(1,ndet)*det_dist )) # x dimension of detector array locations
det_r <- 0.254 # set the radius of the detector face in cm
```

Define energy bins of gamma-ray detectors from minimum energy of backscattered gamma-ray (180 deg scatter) to incident gamma-ray energy.  For initial calculations space the detectors evenly along the x-axis with the source.  Use a bin width that gives 2 bins for the energy resolution of detector (0.4 keV).

```{r}
Eres <- 0.4 # gamma-rEay detector energy resolution in keV
chnwidth <- Eres/2 # assign channel width to give 2 channels for energy resolution of detectors
Emin <- floor(E0/(1+(E0/moc2)*(2))) # set the minimum energy for the detector below E0 back scattered energy
Emax <- floor(E0)
nchns <- (Emax-Emin)/chnwidth
Echn <- tibble ( seq(0,nchns-1)*chnwidth + Emin + chnwidth/2 ) # calculate energy of detector channels
```

Define field of view set by collimator

```{r}
mcoll1 <- 1 # slope of field of view limit 1 from collimator
bcoll1 <- 0 # intercept of field of view limit 1 from collimator
mcoll2 <- -1 # slope of field of view limit 1 from collimator
bcoll1 <- 0 # intercept of field of view limit 1 from collimator
```

Define count times at each position

```{r}
Count <- tibble( times = seq(0,NS-1) ) #set up dataframe for count times
# Count$times <- 1800 #set the count times as a default of 30 m (1800 s)
Count_time <- 3600 #set the count time as a default of 60 m (3600 s)
```


Define object to be imaged

```{r}
Image_loc <- tibble( loc = (seq(0,NX-1)+0.5
                              )*pwidth ) #set up data frame for image plane locations
Count_time <- 3600 #set the count times as a default of 1 hr (3600 s)
# Image_plane <- Matrix (0.00E00, NX, NZ, sparse=TRUE) #initialize the NX by NZ pixel image plane with zeroes
Image_plane <- matrix (0.00E00, NX, NZ,) #initialize the NX by NZ pixel image plane with zeroes
Na <- 6.023E+23 # Avogadro number

Al <- data.frame(massdens = 2.7) # mass density of aluminum in g/cm^3 for object
Al$AW <- 26.98 # atomic weight of aluminum
Al$Z <- 13 # atomic number of aluminum
Al$adens <- Al$massdens*Na/Al$AW
Al$edens <- Al$adens*Al$Z

Image_plane[seq(NX/2,NX/2+1),seq(NZ/2,NZ/2+1)] <- Al$edens # Make four pixels at the middle of the plane for the image

X <- data.frame(X = (seq(0,NX-1)+0.5
                              )*pwidth )
X$Z <- X$X

image( Image_plane )
```

For each nonzero pixel in image plane, compute position (in cm) scattering angle (in degrees), its cosine and the energy E' (in keV) of the scattered gamma ray

```{r}
Ep <- array (0.00E00, dim = c(NX,NZ,ndet) ) #initialize scattered gamma-ray energy in a 3D array of image plane for each detector
Ep_chn <- array (0.00E00, dim = c(NX,NZ,ndet) ) #initialize scattered gamma-ray energy bin in a 3D array of image plane for each detector
Int_prob <- array (0.00E00, dim = c(NX,NZ,ndet) ) #initialize interaction probability in a 3D array of image plane for each detector in (b/Sr)
Events <- array (0.00E00, dim = c(NX,NZ,ndet) ) #initialize interaction probability in a 3D array of image plane for each detector in (b/Sr)

for (i in 1:NX){
  for (j in 1:NZ){

    
    if (Image_plane[i,j] > 0) {
      
      sp_dist <- sqrt(X$X[i]^2 + X$Z[j]^2) # distance from source to pixel in cm
      p_SA <- X$Z[j]*pwidth^2/sp_dist^3    # solid angle of pixel from source in Sr
      
      for (k in 1:ndet){

        pd_dist <- sqrt((det_x[k]-X$X[i])^2 + X$Z[j]^2) # distance from pixel to detector in cm
        cos_theta <- (X$X[i]*(det_x[k]-X$X[i])-X$Z[j]^2) /                                                               (sp_dist*pd_dist) # cosines of angles between vector from source to
                                          #  pixel and pixel to detector for scatter angle
      
        Ep[i,j,k] <- E0/(1+(E0/moc2)*(1-cos_theta))
        Ep_chn[i,j,k] <- floor((Ep[i,j,k]-Emin)/chnwidth)+1
      
        d_SA <- pi*det_r^2*X$Z[j]/pd_dist^3  # solid angle of detector relative to pixel
        
        KN_xs <- (Al$Z*r02/2)*(Ep[i,j,k]/E0)^2*(Ep[i,j,k]/E0 + E0/Ep[i,j,k]                                         - (1-cos_theta^2)) # Klein-Nishina scattering cross section from
                                      # source through pixel into detector
        Int_prob[i,j,k] <- pwidth*Al$edens*KN_xs*1e-24*p_SA                                                                  *d_SA*sp_dist/(4*pi*X$Z[j]) # Scattering probability from
                                      # source through pixel into detector
        Events[i,j,k] <- Int_prob[i,j,k] * Em_rate*Count_time
      
      }
    }
  }
}
```

Now sum the events from each pixel into each detector energy bin to form a "Scattergram" that represents data the Compton scatter tomography system records from the imaged object.

```{r}
# Ep <- array (0.00E00, dim = c(NX,NZ,ndet) ) #initialize scattered gamma-ray energy in a 3D array of image plane for each detector
# Ep_chn <- array (0.00E00, dim = c(NX,NZ,ndet) ) #initialize scattered gamma-ray energy bin in a 3D array of image plane for each detector
# Int_prob <- array (0.00E00, dim = c(NX,NZ,ndet) ) #initialize interaction probability in a 3D array of image plane for each detector in (b/Sr)
# Events <- array (0.00E00, dim = c(NX,NZ,ndet) ) #initialize interaction probability in a 3D array of image plane for each detector in (b/Sr)

ScatGram <- data.frame( Energy = seq(0,nchns-1)*chnwidth + Emin + chnwidth/2)

# chn_counts <- seq(1,nchns)*0 # initiate a vector for the detector channel counts to zero

for (k in 1:ndet){
  chn_counts <- seq(1,nchns)*0 # initiate a vector for the detector channel counts to zero
  
  for (i in 1:NX){

    # sp_dist <- sqrt(X$X[i]^2 + X$Z[j]^2) # distance from source to pixel in cm
    # p_SA <- X$Z[j]*pwidth^2/sp_dist^3    # solid angle of pixel from source in Sr
      
    for (j in 1:NZ){
      if (Image_plane[i,j] > 0) {
        chn <- Ep_chn[i,j,k]
        chn_counts[chn] <- chn_counts[chn] + Events[i,j,k]

        #pd_dist <- sqrt((det_x[k]-X$X[i])^2 + X$Z[j]^2) # distance from pixel to detector in cm
        #cos_theta <- (X$X[i]*(det_x[k]-X$X[i])-X$Z[j]^2) /                                                               (sp_dist*pd_dist) # cosines of angles between vector from source to
                                          #  pixel and pixel to detector for scatter angle
      
        #Ep[i,j,k] <- E0/(1+(E0/moc2)*(1-cos_theta))
        #Ep_chn[i,j,k] <- floor((Ep[i,j,k]-Emin)/chnwidth)+1
      
        #d_SA <- pi*det_r^2*X$Z[j]/pd_dist^3  # solid angle of detector relative to pixel
        
        #KN_xs <- (Al$Z*r02/2)*(Ep[i,j,k]/E0)^2*(Ep[i,j,k]/E0 + E0/Ep[i,j,k]                                         - (1-cos_theta^2)) # Klein-Nishina scattering cross section from
                                      # source through pixel into detector
        #Int_prob[i,j,k] <- pwidth*Al$edens*KN_xs*1e-24*p_SA                                                                  *d_SA*sp_dist/(4*pi*X$Z[j]) # Scattering probability from
                                      # source through pixel into detector
        #Events[i,j,k] <- Int_prob[i,j,k] * Em_rate*Count$times[k]
      
      }
    }
  }
  col <- as.character(det_x[k])
  ScatGram[col] <- chn_counts
}

```

plot the scattergram for the last detector

```{r}
plot(ScatGram$Energy, ScatGram[,13])

```

```{r}
ScatterMat <- ScatGram %>%
  as.matrix()

ScatterImg <- ScatterMat %>%
  image()

write_csv(ScatGram, "ScatterMat.csv")
```

