---
title: "ANN"
author: "Dominic Cugliari"
date: "2024-05-30"
output: html_document
---

Create working directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath(".."))
```

Load libraries

```{r}
library(neuralnet)
library(tidyverse)
library(binhf)
library(Matrix)
library(ggplot2)
library(pracma)
library(OpenImageR)
```

read in .csv of reference image and pass reference pixel value/location into a 'output' vector and 'x' and 'z' vectors.

It is important to note that all the "zeros" have the value 0.001. I don't know how this affects the training, but it would be worth getting a new training set with real zeroes and trying.

I have been working to figure out how to train this and have the error print iteratively so we can see if it is improving. I have not figured this out yet.

```{r}
# read in reference data
m <- read.csv("raw_data/65ref.csv")
m <- m[, -c(1)]
v <- matrix(nrow = 3, ncol = 0)
# Populate 'v' with columns, rows, and values from 'm'
for (i in 1:nrow(m)){
  for (j in 1:ncol(m)){
    v <- cbind(v, c(i,j,m[i,j]))
  }
}

# Create empty vectors `x`, `z`, and 'output'
x <- vector(length = 0) 
z <- vector(length = 0)
output <- vector(length = 0)

# Extract values from each row of `v`
for (i in 1:ncol(v)){ 
     x <- c(x, v[2, i])
     z <- c(z, v[1, i])
     output <- c(output, v[3, i]) 
   }
x <- (x-(max(x)+1)/2) # make x and z correct for the shift flip sum code
z <- (z-(max(z)+1)/2) # origin referenced coordinates
```

Train neural net

Hello World

```{r}
# read in sinogram data
sino <- read.csv('raw_data/65sin.csv')
sino <- sino[, -c(1)]
# number of projections
nAngle  <- ncol(sino)
# number of detectors
nDetect <- nrow(sino)
# Degrees between projections
angleStep <- 180/(nAngle-1) 
# compute angle of each projection in sinogram
degree <- data.frame(angle = seq(0,nAngle-1)*angleStep) 

# initialize weights
nn <- list()
nn$weights <- list()

nn$weights[[1]] <- matrix(0, nrow = 16, ncol = 2)
nn$weights[[2]] <- matrix(0, nrow = 3, ncol = 1)
error <- matrix(nrow = 1, ncol = 0)
for (i in 1:length(output)){
   # compute shift of sinogram detector positions for each projection based on current image pixel
   degree$pixel <- round(x[i]*cos(pi*degree$angle/180) + z[i]*sin(pi*degree$angle/180))
   # transpose the sinogram
   tsino <- t(sino)
   # create a space with dimension tsino
   ssino <- tsino   

   ######################################################## SHIFT SUM FLIP 
   # Perform shift according to dShift
   for (j in 1:nAngle){ 
      shift_amount <- degree$pixel[j]
      shifted_row <- circshift(tsino[j, ], shift_amount)
      ssino[j, ] <- shifted_row
      }
   # perform summation
   f_vector <- colSums(ssino) 
   # perform flip
   z_vector <- rev(f_vector)  
   
   ################################################################## DATA
   data <- t(as_tibble(c(z_vector, output[i])))
   # Define names of the columns for the data to used by function 
   ANN_names <- c(paste0("z",(as.character(c(seq(1:nDetect))))),"o") 
  
    colnames(data) <- ANN_names
   ############################################################### FORMULA
   f <- as.formula(paste(ANN_names[nDetect+1],
                         "~",
                         paste(ANN_names[!ANN_names %in% "o"], collapse = " + ")
                         )
                   ) 
   ############################################################ NEURAL NET
   nn <-neuralnet(f, data, hidden = 2,
                  startweights = nn$weights)
   error[i] <- nn$result.matrix[1]
   }

plot(nn, rep = "best")
# Creating matrices 'm1' which contains the weights from the first layer and 'm2' which contains the
# weights from the second layer
m1 <- nn$weights[[1]][[1]]
m2 <- nn$weights[[1]][[2]]
m1 <- m1 %>% as.data.frame()
m2 <- m2 %>% as.data.frame()
```

```{r}
filt_sinoA <- sino
filt_sinoB <- sino
#number_cols <- ncol(sino)

#for(val in 1:number_cols){
                   # filt_sinoA[,val] <- convolve(sino[,val], m1[2:91,1], type ="circular")
#}

#filt_sinoA %>% 
 # as.matrix()

#write_csv(filt_sinoA, "raw_data/filtsinoa-DC.csv")


#for(val in 1:number_cols){
            #       filt_sinoB[,val] <- convolve(sino[,val], m1[2:91,2], type ="circular")
#}

#filt_sinoB %>% 
#  as.matrix()

#write_csv(filt_sinoB, "raw_data/filtsinob-DC.csv")


write_csv(m1, "raw_data/m1-DC.csv")
write_csv(m2, "raw_data/m2-DC.csv")


```

```{r}
filt_sinoAT <- t(filt_sinoA)
w1 <- m1[2:91,1] %>% as.vector()

result1 <- filt_sinoAT * w1
result1 <- t(result1)
result1 <- result1 %>% as.data.frame()

write_csv(result1, "raw_data/filtsinoa-DC.csv")


filt_sinoBT <- t(filt_sinoB)
w2 <- m1[2:91,2] %>% as.vector()

result2 <- filt_sinoBT * w2
result2 <- t(result2)
result2 <- result2 %>% as.data.frame()

write_csv(result2, "raw_data/filtsinob-DC.csv")

```


Read in the FBP image!!!

```{r}
 p <- read.csv("C:/Users/capa01/Documents/R-Projects/ANN_work/clean_data/FBP1-DC.csv")
p <- p %>% as.matrix()
image(p)
```







