---
title: "Palmer Penguins Neural Net"
author: "Dominic Cugliari"
date: "2024-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r}
library(torch)
library(tidymodels)
library(palmerpenguins)
library(dplyr)
```


Let's glimpse the data.

```{r}
penguins %>% glimpse()
```

We will be predicting species using a few of the variables.  To build the data set for torch, we can run this code:

```{r}
penguins_dataset <- dataset(
  name = "penguins_dataset()",
  initialize = function(df) {
    df <- na.omit(df)
    self$x <- as.matrix(df[, 3:6]) %>% torch_tensor()
    self$y <- torch_tensor(
      as.numeric(df$species)
    )$to(torch_long())
  },
  .getitem = function(i) {
    list(x = self$x[i, ], y = self$y[i])
  },
  .length = function() {
    dim(self$x)[1]
  }
)

# Check a few things to make sure it worked

ds <- penguins_dataset(penguins)
length(ds)
ds[1:10]
```

When you already have a tensor around, or something that’s readily converted to one, you can make use of a built-in dataset() generator: tensor_dataset(). This function can be passed any number of tensors; each batch item then is a list of tensor values:

```{r}
three <- tensor_dataset(
  torch_randn(10), torch_randn(10), torch_randn(10)
)
three[1]
```

In our Penguins data set, we only need two lines of code.

```{r}
penguins <- na.omit(penguins)
ds <- tensor_dataset(
  torch_tensor(as.matrix(penguins[, 3:6])),
  torch_tensor(
    as.numeric(penguins$species)
  )$to(torch_long())
)

ds[1:10]
```

Thirdly and finally, here is the most effortless possible way.

```{r}
library(torchvision)
library(luz)

dir <- "~/.torch-datasets"

ds <- mnist_dataset(
  root = dir,
  train = TRUE, # default
  download = TRUE,
  transform = function(x) {
    x %>% transform_to_tensor() 
  }
)

first <- ds[1]
cat("Image shape: ", first$x$shape, " Label: ", first$y, "\n")
```


```{r}
dl <- dataloader(ds, batch_size = 32, shuffle = TRUE)
length(dl)
first_batch <- dl %>%
  # obtain an iterator for this dataloader
  dataloader_make_iter() %>% 
  dataloader_next()

dim(first_batch$x)
dim(first_batch$y)
```


```{r}
# input dimensionality (number of input features)
d_in <- 3
# number of observations in training set
n <- 1000

x <- torch_randn(n, d_in)
coefs <- c(0.2, -1.3, -0.5)
y <- x$matmul(coefs)$unsqueeze(2) + torch_randn(n, 1)

ds <- tensor_dataset(x, y)

dl <- dataloader(ds, batch_size = 100, shuffle = TRUE)
```


```{r}
# dimensionality of hidden layer
d_hidden <- 32
# output dimensionality (number of predicted features)
d_out <- 1

net <- nn_module(
  initialize = function(d_in, d_hidden, d_out) {
    self$net <- nn_sequential(
      nn_linear(d_in, d_hidden),
      nn_relu(),
      nn_linear(d_hidden, d_out)
    )
  },
  forward = function(x) {
    self$net(x)
  }
)
```



```{r}
fitted <- net %>%
  setup(loss = nn_mse_loss(), optimizer = optim_adam) %>%
  set_hparams(
    d_in = d_in,
    d_hidden = d_hidden, d_out = d_out
  ) %>%
  fit(dl, epochs = 200)
```

```{r}
fitted <- net %>%
  setup(loss = nn_mse_loss(), optimizer = optim_adam) %>%
  set_hparams(
    d_in = d_in,
    d_hidden = d_hidden, d_out = d_out
  ) %>%
  fit(ds, epochs = 200)
```

```{r}
fitted <- net %>%
  setup(loss = nn_mse_loss(), optimizer = optim_adam) %>%
  set_hparams(
    d_in = d_in,
    d_hidden = d_hidden, d_out = d_out
  ) %>%
  fit(ds, epochs = 200)
```

```{r}
fitted <- net %>%
  setup(loss = nn_mse_loss(), optimizer = optim_adam) %>%
  set_hparams(
    d_in = d_in,
    d_hidden = d_hidden, d_out = d_out
  ) %>%
  fit(list(x, y), epochs = 200)
```

```{r}
fitted <- net %>%
  setup(loss = nn_mse_loss(), optimizer = optim_adam) %>%
  set_hparams(
    d_in = d_in,
    d_hidden = d_hidden, d_out = d_out
  ) %>%
  fit(list(as.matrix(x), as.matrix(y)), epochs = 200)


```

```{r}
fitted <- net %>%
  setup(
    loss = nn_mse_loss(),
    optimizer = optim_adam,
    metrics = list(luz_metric_mae())
  ) %>%
  fit(...)
```



```{r}
train_ids <- sample(1:length(ds), size = 0.6 * length(ds))
valid_ids <- sample(
  setdiff(1:length(ds), train_ids),
  size = 0.2 * length(ds)
)
test_ids <- setdiff(
  1:length(ds),
  union(train_ids, valid_ids)
)

train_ds <- dataset_subset(ds, indices = train_ids)
valid_ds <- dataset_subset(ds, indices = valid_ids)
test_ds <- dataset_subset(ds, indices = test_ids)

train_dl <- dataloader(train_ds,
  batch_size = 100, shuffle = TRUE
)
valid_dl <- dataloader(valid_ds, batch_size = 100)
test_dl <- dataloader(test_ds, batch_size = 100)
```





















